### 2016-07-12 cross_compilation.tex
Toolchain description done

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and / or compilation time. Thus, in
this Section we present a procedure of how to cross-compile C++
source code from the testbed administrator \ac{PC} for the \ac{ARM}
architecture of the \ac{Raspi}s. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator \ac{PC} in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program to run
on a \ac{Raspi}.

@@

%The toolchain is mandatory on most consumer \ac{PC}s due to the different
%processor architecture.
%
%Furthermore, we will cover how to cross compile a simple C++ example and
%Kodo-cpp as well as how the binaries can be copied from a \ac{PC} to a
%\ac{Raspi} and executed.

\subsection{Setup toolchain on PC}

%Besides the previous description (\textbf{Include compiling Kodo from the
%RasPi from the scratch}), the testbed administrator can compile Kodo in its
%personal workstation and transfer the generated binaries directly to
%a path in the \ac{Raspi}. To achieve this, we get a toolchain that
%contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
%for the \ac{Raspi}. Therefore, we strongly recommend any testbed
%administrator to do the following procedure. In what follows, we provide
%the instructions considering that the NFS server uses the \texttt{\$HOME}
%directory as the working directory. However, the administrator may choose
%some other working directory of its preference if desired.

\item Create toolchain directory:
\begin{lstlisting}[]
$ TOOLCHAINDIR="${HOME}/toolchains"
$ mkdir -p $TOOLCHAINDIR
$ cd $TOOLCHAINDIR
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Download the \ac{Raspi} toolchain for 64-bit Linux from.
A more recent \ac{Raspi} toolchain may be available at
\texttt{http://buildbot.steinwurf.dk/toolchains/linux/}:

\begin{lstlisting}[]
$ TOOLCHAIN="raspberry-gxx49-arm"
$ wget http://kom.aau.dk/project/TuneSCode/raspi/${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Extract the downloaded file:
%locally in the NFS server. After
%this operation, there should be a new directory for the toolchain
%in the server. \\
\begin{lstlisting}[]
$ unzip ${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item (Optional) Delete zip file:
%locally in the NFS server. After
%this operation, there should be a new directory for the toolchain
%in the server. \\
\begin{lstlisting}[]
$ rm ${TOOLCHAIN}.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item We can now verify that the ARM cross compiler is working:

\begin{lstlisting}[]
$ ${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++ --version
raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
Copyright (C) 2014 Free Software Foundation, Inc.
This is free software; see the source for copying conditions.  There is NO
warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\begin{lstlisting}[]
%$ echo "${TOOLCHAINDIR}/${TOOLCHAIN}/bin/${TOOLCHAIN}-g++"
%/home/<USER>/toolchains/raspberry-gxx49-arm/bin/raspberry-gxx49-arm-g++
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

%    The ARM cross compilers should now be located in

\item Make toolchain binaries available systemwide. Instead of calling
the ARM cross compiler using its full path, we can make it accessible
from the command shell systemwide. One way to do this is by adding the
toolchain binaries directory to the Linux environment variable \texttt{PATH}
when the \ac{OS} starts up:

%\item Add toolchain binaries to \texttt{PATH}. Instead specifying the full
%path to the toolchain binaries we can instead tell the operation system
%where to search for it. This makes the toolchain binaries available
%systemwide.

\begin{lstlisting}[]
$ echo PATH=\"\$PATH:${TOOLCHAINDIR}/${TOOLCHAIN}/bin\" >> ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
%$ printf 'PATH="%s/%s/bin"\n' "${TOOLCHAINDIR}" "${TOOLCHAIN}" >> ${HOME}/.profile

%\item Add the \texttt{bin} folder of the toolchain to the \texttt{PATH}
%Linux environment variable of the server. This will help the server OS
%to recognize the location of the compiler command, which will be needed
%later. To do so, edit the \texttt{\$HOME/.profile} to add in a newline:
%\texttt{PATH="\$PATH:\$HOME/raspberry-gxx49-arm/bin"}. Save the
%\texttt{\$HOME/.profile}. \\

\texttt{.profile} should now contain the line we inserted. There may be more
code in you file.
% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$HOME/.profile}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
PATH="$PATH:/home/<USERNAME>/toolchains/raspberry-gxx49-arm/bin"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item Update \texttt{PATH}. Source \texttt{.profile} to make the changes
take effect in your system:
\begin{lstlisting}[]
$ source ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\item ARM cross compiler should now be available systemwide:

\begin{lstlisting}[]
$ raspberry-gxx49-arm-g++ --version
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\item Restart the server session in order for the changes made in the
%previous step take effect. To verify this, open a new terminal and type:
%\texttt{raspberry-gxx49-arm-g++ --version}. A correct binary installation
%should return an output similar to:
%
%\texttt{raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
%Copyright (C) 2014 Free Software Foundation, Inc.
%This is free software; see the source for copying conditions.  There is NO
%warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.} \\

>>>>

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and / or compilation time. Thus, in
this Section we present a procedure of how to cross-compile C++
source code from the testbed administrator \ac{PC} for the \ac{ARM}
architecture of the \ac{Raspi}s. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator \ac{PC} in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program and
copying the generated binaries with \ac{SSH} to run locally
on a \ac{Raspi}.

@@

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain in its \ac{PC}.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from / to.

Given that compiling a toolchaing can be an ardous task, we get the toolchain
recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
This toolchain is available from \cite{2016rpitoolchain} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$TOOLCHAIN} directory as the working directory. However,
the testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

\begin{lstlisting}[]
$ export TOOLCHAINDIR="${HOME}/toolchains"
$ mkdir -p $TOOLCHAINDIR
$ cd $TOOLCHAINDIR
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Later, we download the \ac{Raspi} toolchain for 64-bit Linux. However,
instead of working with the binaries from \cite{2016rpitoolchain},
we use a zipped version with the original binaries and renamed symbolic
links to the original files available in
\cite{2016steinwurfrpitoolchain}. Finally, we unzip the downloaded file.
This is made as follows:

\begin{lstlisting}[]
$ wget http://files.steinwurf.com/toolchains/linux/raspberry-gcc-4.9.3/raspberry-gxx493-arm.zip
$ unzip raspberry-gxx493-arm.zip
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Instead of calling the ARM cross compiler using its full path, we make
the binaries accessible from the command shell systemwide. A way to do
this is by adding the following commands in the \texttt{\${HOME}/.profile}
as follows:

\begin{lstlisting}[]
$ sed -i '$a export TOOLCHAINDIR=\"$HOME/toolchains\"' ${HOME}/.profile
$ sed -i '$a export TOOLCHAINBINARY=\"raspberry-gxx49-arm-g++\"' ${HOME}/.profile
$ sed -i '$a PATH=\"\$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin\"' ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This helps the \ac{OS} to recognize the location of the compiler
command everytime a new shell is opened. The \texttt{.profile} should now
contain the lines we inserted. There might be more other
code in your file. Leave other parts unmodified.
% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$HOME/.profile}@>
<@\textcolor{gray}{
---------------------------}
\Reactivatenumber @>
...
export TOOLCHAINDIR="$HOME/toolchains"
export TOOLCHAINBINARY="raspberry-gxx49-arm-g++"
PATH="$PATH:${TOOLCHAINDIR}/arm-rpi-4.9.3-linux-gnueabihf/bin"
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To update the \texttt{PATH} variable and the \texttt{.profile}, we use the
\texttt{source} command for the changes take effect in the administrator
system:
\begin{lstlisting}[]
$ source ${HOME}/.profile
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

A working ARM cross-compiler in the testbed administrator \ac{PC} should
output the following:

\begin{lstlisting}[]
$ ${TOOLCHAINBINARY} --version
raspberry-gxx49-arm-g++ (crosstool-NG crosstool-ng-1.22.0-88-g8460611) 4.9.3
Copyright (C) 2015 Free Software Foundation, Inc.
This is free software; see the source for copying conditions.  There is NO
warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
